Semi-Parametric Manifold Clustering

### 2022-01-12 curve-fitting-methods/curve-fitting-scratch.Rmd
added example

library(ggplot2)
import::from(magrittr, `%>%`)
```

@@

```{r}
source('~/dev/pabm-grdpg/functions.R')
source('~/dev/dissertation/curve-fitting-methods/functions.R')
```

@@

It can be shown that each iteration decreases the objective unless we are at a stationary point.

## Model-Based Clustering

The CLT property of the ASE may allow us to approximate $W x - \hat{x} \sim \mathcal{N}(f_k(t), \Sigma_k)$.
One possibly reasonable assumption for $\Sigma_k$ is $\Sigma_k = \Sigma_k(t) =  \diag(\sigma_{k1}^2(t), ..., \sigma_{kd}^2(t))$. A plot of the squared errors of cluster 1 in the previous example reveals the following:

```{r, fig.height = 2, fig.width = 8}
theta1 <- manifold.out$theta[[1]]
theta2 <- manifold.out$theta[[2]]
t. <- manifold.out$t
X1 <- param.curve.quadr.2(t.[[1]], 
                          theta1[1], theta1[2], theta1[3],
                          theta1[4], theta1[5], theta1[6])
X2 <- param.curve.quadr.2(t.[[2]], 
                          theta2[1], theta2[2], theta2[3],
                          theta2[4], theta2[5], theta2[6])
# plot(Xhat[zhat.manifold == 1, 1] - X1[, 1])
# plot(Xhat[zhat.manifold == 1, 2] - X1[, 2])

p1 <- ggplot() + 
  geom_point(aes(x = t.[[1]], y = (Xhat[zhat.manifold == 1, 1] - X1[, 1]) ^ 2)) + 
  stat_smooth(aes(x = t.[[1]], y = (Xhat[zhat.manifold == 1, 1] - X1[, 1]) ^ 2)) + 
  theme_bw() + 
  labs(x = 't', y = expression((x[1] - hat(x)[1])^2))

p2 <- ggplot() + 
  geom_point(aes(x = t.[[1]], y = (Xhat[zhat.manifold == 1, 2] - X1[, 2]) ^ 2)) + 
  stat_smooth(aes(x = t.[[1]], y = (Xhat[zhat.manifold == 1, 2] - X1[, 2]) ^ 2)) + 
  theme_bw() + 
  labs(x = 't', y = expression((x[2] - hat(x)[2])^2))

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

Fitting some curve to each $\sigma_{kr}^2(t)$ allows us to compute estimated probabilities of each point belonging to each cluster/manifold: $\hat{q}_{ik} = \mathcal{N}(x_i \mid g_k(\hat{t}_i), \Sigma_k(\hat{t}_i))$. 
Then we can use weighted least squares to estimate the parameters of $g$:

$$\hat{p}_k = (T^\top Q_k T)^{-1} T^\top Q_k X$$

where $Q_k = \diag(q_{1k}, ..., q_{nk})$.

>>>>

library(ggplot2)
import::from(magrittr, `%>%`, `%<>%`)
```

@@

```{r}
source('~/dev/pabm-grdpg/functions.R')
source('~/dev/dissertation/curve-fitting-methods/functions.R')
source('https://mtrosset.pages.iu.edu/Courses/675/graph.r')
```

@@

**TBD**: It can be shown that each iteration decreases the objective unless we are at a stationary point.

## Harry Potter Emnity Graph

\begin{example}
In this section, we analyze the Harry Potter emnity graph. 
Rubin-Delanchy et al. suggested applying the GRDPG ASE to this graph with $p = 1$ and $q = 1$, revealing a DCBM-like structure in the latent space. 
Passino and Heard instead fit Gaussian processes to the embedding. 
Here, we consider what happens when we apply the Bezier curve clustering method. 
\end{example}

```{r, fig.width = 8}
relations.df <- readr::read_csv('~/dev/potter-network/data/relations.csv') %>% 
  dplyr::mutate(source = source + 1, target = target + 1)
characters.df <- readr::read_csv('~/dev/potter-network/data/characters.csv') %>% 
  dplyr::mutate(id = id + 1)
n <- nrow(characters.df)

A <- matrix(0, nrow = n, ncol = n)
rownames(A) <- colnames(A) <- characters.df$name
for (i in seq(nrow(relations.df))) {
  if (relations.df$type[i] == '-') {
    A[relations.df$source[i], relations.df$target[i]] %<>% magrittr::add(1)
  } else {
    A[relations.df$source[i], relations.df$target[i]] %<>% magrittr::subtract(1)
  }
}
A[A < 0] <- 0
A <- A + t(A)
A <- sign(A)
diag(A) <- 0
deg.A <- colSums(A)

A <- A[deg.A > 0, deg.A > 0]

A.eigen <- eigen(A, symmetric = TRUE)
Xhat <- embedding(A, p = 1, q = 1)

# ggplot() + 
#   geom_point(aes(x = Xhat[, 1], y = Xhat[, 2])) + 
#   ggrepel::geom_text_repel(aes(x = Xhat[, 1], y = Xhat[, 2], 
#                                label = characters.df$name[deg.A > 0]))

z.init <- kmeans(sweep(Xhat, 1, apply(Xhat, 1, function(x) sqrt(sum(x ^ 2))), `/`), 2)$cluster
```

```{r, fig.width = 10, fig.height = 4, cache = TRUE}
manifold.out <- manifold.clustering.quadr.2(A, 2, p = 1, q = 1, 
                                            initialization = z.init, 
                                            eps.isomap = .7,
                                            method = 'bezier')

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(zhat.manifold))) + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  theme(legend.position = 'none') +
  geom_path(aes(x = Y[, 1], y = Y[, 2],
                group = factor(z.))) +
  ggrepel::geom_text_repel(aes(x = Xhat[, 1], y = Xhat[, 2],
                               label = characters.df$name[deg.A > 0])) + 
  xlim(min(Xhat[, 1]), max(Xhat[, 1])) + ylim(min(Xhat[, 2]), max(Xhat[, 2]))
```

### 2022-01-12 curve-fitting-methods/curve-fitting-scratch.tex
added example

It can be shown that each iteration decreases the objective unless we
are at a stationary point.

\hypertarget{model-based-clustering}{%
\subsection{Model-Based Clustering}\label{model-based-clustering}}

The CLT property of the ASE may allow us to approximate
\(W x - \hat{x} \sim \mathcal{N}(f_k(t), \Sigma_k)\). One possibly
reasonable assumption for \(\Sigma_k\) is
\(\Sigma_k = \Sigma_k(t) = \mathrm{diag}(\sigma_{k1}^2(t), ..., \sigma_{kd}^2(t))\).
A plot of the squared errors of cluster 1 in the previous example
reveals the following:

\begin{center}\includegraphics{curve-fitting-scratch_files/figure-latex/unnamed-chunk-10-1} \end{center}

Fitting some curve to each \(\sigma_{kr}^2(t)\) allows us to compute
estimated probabilities of each point belonging to each
cluster/manifold:
\(\hat{q}_{ik} = \mathcal{N}(x_i \mid g_k(\hat{t}_i), \Sigma_k(\hat{t}_i))\).
Then we can use weighted least squares to estimate the parameters of
\(g\):

\[\hat{p}_k = (T^\top Q_k T)^{-1} T^\top Q_k X\]

where \(Q_k = \mathrm{diag}(q_{1k}, ..., q_{nk})\).

>>>>

\textbf{TBD}: It can be shown that each iteration decreases the
objective unless we are at a stationary point.

\hypertarget{harry-potter-emnity-graph}{%
\subsection{Harry Potter Emnity Graph}\label{harry-potter-emnity-graph}}

\begin{example}
In this section, we analyze the Harry Potter emnity graph. 
Rubin-Delanchy et al. suggested applying the GRDPG ASE to this graph with $p = 1$ and $q = 1$, revealing a DCBM-like structure in the latent space. 
Passino and Heard instead fit Gaussian processes to the embedding. 
Here, we consider what happens when we apply the Bezier curve clustering method. 
\end{example}

\begin{center}\includegraphics{curve-fitting-scratch_files/figure-latex/unnamed-chunk-11-1} \end{center}
